---
title: "Holmquist et al. 2024 manuscript figures"
author: "Holmquist, AJ"
date: "2024-09-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
knitr::opts_knit$set(root.dir = here())
```

Run all code from primary analysis code prior to reproducing the below figures
```{r}
# Figures
library(tidyverse)
library(ggpubr)
library(pROC)
library(DHARMa)
library(ggsignif)
```

Figure 2. Interaction between NaCl molarity and collection year in prediction 
of DNA concentration, 260/230 and 260/280 ratios NaCl figures
```{r}
nacl_palette <- c("#F7B538", "#DB7C26", "#DD4B1A")

# Figure
plot_260.230 <- 
  nacl_filt %>%
  mutate(
    nacl = factor(nacl, 
                  levels = c("1M", "2M", "2.5M"))) %>%
  ggplot(
       aes(year, r_260.230_norm)) +
  geom_point(aes(color = nacl), size = 2, alpha = 0.3) +
  geom_smooth(aes(color = nacl, linetype = nacl), method = "lm", se = F) +
  theme_minimal() +
  scale_color_manual(values = nacl_palette) +
  xlab("Collection year") +
  ylab("Log(260/230)") +
  labs(color = "NaCl molarity", linetype = "NaCl molarity")

plot_260.280 <- 
  nacl_filt %>%
  mutate(nacl = factor(nacl, levels = c("1M", "2M", "2.5M"))) %>%
  ggplot(aes(year, r_260.280_norm)) +
  geom_point(aes(color = nacl), size = 2, alpha = 0.3) +
  geom_smooth(aes(color = nacl, linetype = nacl), method = "lm", se = F) +
  theme_minimal() +
  scale_color_manual(values = nacl_palette) +
  xlab("Collection year") +
  ylab("Log(260/280)") +
  labs(color = "NaCl molarity", linetype = "NaCl molarity")

conc_plot <- 
  nacl_filt %>%
  mutate(nacl = factor(nacl, levels = c("1M", "2M", "2.5M"))) %>%
  ggplot(aes(year, conc_norm)) +
  geom_point(aes(color = nacl), size = 2, alpha = 0.3) +
  geom_smooth(aes(color = nacl, linetype = nacl), method = "lm", se = F) +
  theme_minimal() +
  scale_color_manual(values = nacl_palette) +
  xlab("Collection year") +
  ylab("Log(DNA concentration)") +
  labs(color = "NaCl molarity", linetype = "NaCl molarity")

interaction_nacl_year <- 
  ggarrange(conc_plot, plot_260.230, plot_260.280, common.legend = T, 
            nrow = 1, ncol = 3)

ggsave("figure2.png", 
       width = 9, height = 4, units = "in", 
       dpi = 500)
```

Figure 3. Amplification success of the 658 bp full COI gene and the 181 bp 
mini-barcode using different bead ratios, PEG-8000 concentrations, and NaCl molarities
```{r}
amplification_success_bar <- 
  nacl_filt %>%
  pivot_longer(cols = c(mini, full),
               names_to = "amplicon", 
               values_to = "success") %>%
  mutate(success = ifelse(success == 1, "Yes", "No"),
         nacl = factor(nacl, levels = c("1M", "2M", "2.5M")),
         amplicon = ifelse(amplicon == "full", 
                           "Full barcode (658bp)",
                           "Mini barcode (181bp)")) %>%
  ggplot() +
  geom_bar(aes(nacl, fill = success),
           linewidth = 2, alpha = 0.9) +
  facet_grid(peg~ratio) +
  theme_minimal() +
  scale_fill_manual(values = c("#0E2D2F", "#7399AB")) +
  scale_color_manual(values = nacl_palette) +
  facet_grid(ratio+peg ~ amplicon) +
  labs(
    fill = "Amplification success"
  ) +
  theme(
    axis.title.y = element_blank(),  # Remove y-axis title
    axis.text.y = element_blank(),   # Remove y-axis text labels
    axis.ticks.y = element_blank(),
    legend.position = "bottom") +
  xlab("NaCl")

ggsave("figures/figure3.png", 
       height = 4.5, width = 4.5, units = "in",
       dpi = 500)

```

Figure 4. Comparison of DNA concentrations produced from the fly samples 
(H. sonomensis) and beetle samples (P. andrewsi) using four different 
extraction methods. 
```{r}
extract_colors <- c("#DFB77C", "#90857A","#597788", "#8D5D2A")

dip_box <- 
  ggplot(diptera, aes(x = method, y = qubit)) +
  geom_boxplot(aes(fill = method)) +
  geom_jitter(alpha = 0.5, size = 2) +
  geom_signif(comparisons = list(c("Qiagen DNeasy", "HotShot"), 
                                 c("Qiagen DNeasy", "Qiagen Puregene"),
                                 c("Qiagen DNeasy", "Bead-based method")),
              annotations = c("***", "***", "***"), 
              y_position = c(2.1, 2.4, 2.7),
              map_signif_level = TRUE) +
  ylim(0, 3) +
  xlab("Extraction method") +
  scale_fill_manual(values = extract_colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, size = 9)) +
  labs(title = "Fly",
       fill = "Extraction method") +
  ylab("Qubit Concentration (ng/uL)")

coleo_box <- ggplot(coleo, aes(x = method, y = qubit)) +
  geom_boxplot(aes(fill = method)) +
  geom_jitter(alpha = 0.5, size = 2) + 
  ylim(0, 32) +
  xlab("Extraction method") +
  geom_signif(comparisons = list(c("Qiagen DNeasy", "HotShot"), 
                                 c("Qiagen DNeasy", "Qiagen Puregene"),
                                 c("Qiagen DNeasy", "Bead-based method")),
              annotations = c("**", "***", "n.s."), 
              y_position = c(23, 26, 29),
              map_signif_level = TRUE) +
  scale_fill_manual(values = extract_colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, size = 9)) +
  labs(title = "Beetle",
       fill = "Extraction method") +
  ylab("Qubit Concentration (ng/uL)")

conc_boxplot <-
  ggarrange(dip_box, coleo_box, common.legend = TRUE)

ggsave("figures/figure4.png", 
       width = 6.5, height = 5, units = "in", 
       dpi = 500)

```

Figure 6. Boxplot showing the impact of input tissue type on normalized DNA concentration. 
```{r}
conc_method_boxplot <- 
  highthroughput_nanodrop %>%
  mutate(method = ifelse(method == "destructive",
                         "Subsampled tissue",
                         "Non-destructive whole-body")) %>%
  ggplot(aes(method, conc_norm)) +
  geom_boxplot(aes(fill = method),
               outlier.size = 2,
               alpha = 0.8) +
  geom_jitter(aes(shape = method), alpha = 0.3) +
  geom_signif(comparisons = list(c("Subsampled tissue", 
                                   "Non-destructive whole-body")),
              annotations = "***", 
              y_position = 7,
              map_signif_level = TRUE)  +
  scale_fill_manual(values = c("#7A5C58", "#8D80AD")) +
  theme_minimal() +
  ylab("Normalized DNA concentration") +
  xlab("") +
  labs(fill = "Lysis method",
       shape = "Lysis method") +
  theme(legend.position = "top")

ggsave("figures/figure6.png", 
       width = 5, height = 4, units = "in",
       dpi = 500)

```

Supporting Materials:

S1. Figure 1
```{r}
norm_conc_box <- 
  nacl %>%
  ggplot(aes(catalog, concentration)) +
  geom_boxplot(aes(fill = catalog), alpha = 0.9) +
  geom_jitter(alpha = 0.1) + 
  theme_minimal() +
  theme(legend.position = "none", 
        axis.text.x = element_blank()) +
   xlab("Sample") +
   ylab("DNA Concentration")

ggsave("supplementary/figures/S1Figure1.png", 
       width = 5.8, height = 4.6, unit = "in", dpi = 500)
```

S1. Figure 2 - Figure 7 (model diagnostics)
```{r}

# Concentration
sim_res_conc <- simulateResiduals(fittedModel = conc_naclyear)
DHARMa::plotQQunif(sim_res_conc)
DHARMa::plotResiduals(sim_res_conc)

# 260/280
sim_res_260.280 <- simulateResiduals(fittedModel = nacl_260.280) 
DHARMa::plotQQunif(sim_res_260.280)
DHARMa::plotResiduals(sim_res_260.280)
      
# 260/230
sim_res_260.230 <- simulateResiduals(fittedModel = naclyear_260.230)
DHARMa::plotQQunif(sim_res_260.230)
DHARMa::plotResiduals(sim_res_260.230)
      
# PCR Model
roc_curve <- 
  roc(response = nacl_filt$mini, predictor = predict(pcr_model_nacl, type = "response"))
plot(roc_curve)

sim_res_pcr <- simulateResiduals(fittedModel = pcr_model_nacl)
DHARMa::plotQQunif(sim_res_pcr)
DHARMa::plotResiduals(sim_res_pcr)
```

S1. Figure 8 
```{r}
ggplot(highthroughput_nanodrop, aes(year_norm, conc_norm)) +
  geom_point(aes(year_norm, conc_norm), alpha = 0.5) +
  geom_smooth(aes(color = method), method = "lm") +
  theme_minimal()

ggsave("supplementary/figures/S1Figure8.png", 
       width = 6, height = 4.5, unit = "in", 
       dpi = 500)
```
